{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-19 thesis/education.tex
add photo of ze studence box plot; rough draft for now; colors need work etc
education prose revise

As described in \sect{\ref{sec:landslide-setup}},
all setup from the user's point of view is handled through the {\tt p2-setup.sh} script%
\footnote{PSU's version is called {\tt psu-setup.sh};
in this section, unless otherwise noted, {\tt p2-setup.sh} will refer to both.
}.
It, its helper scripts (\sect{\ref{sec:landslide-glue}}),
and the {\tt landslide} script itself contain several checks to prevent
studence
from accidentally misusing Landslide in ways that could produce mysterious crashes, false bug reports, and so on
(the need for each one, as the reader might imagine, discovered through bitter experience).
These include:
\begin{itemize}
	\item {\tt p2-setup.sh} checks if the directory argument correctly points at the top-level P2 basecode directory
		rather than any subdirectories such as {\tt user/libthread/}.
	\item {\tt check-need-p2-setup-again.sh} checks if any source files in the original P2 source directory
		(the argument supplied to {\tt p2-setup.sh}),
		in case the student hoped to fix some bug and verify their fix but forgot to re-run the setup script.
	\item {\tt landslide} checks the supplied test name matches one of the endorsed Landslide-friendly tests
		(students love trying to run Landslide with {\tt racer}, {\tt largetest},
		or even the string {\tt OPTIONS}).
	\item {\tt landslide} checks if any other instance of itself is simultaneously running in the same directory,
		and if so, refuses to do so and advises the student
		to {\tt git clone} the repository afresh for simultaneous use%
		\footnote{This is ironically implemented with a non-atomic lock file
		and should really be using {\tt flock} instead.
		}.
\end{itemize}
\vspace{1em}

\noindent Landslide also includes several P2-specific instrumentations and features to cope with various student irregularities:
\begin{itemize}
	\item Quicksand emits different combinations of {\tt within\_function}/{\tt without\_function} directives
		for Landslide depending on the name of the test.
		For example, for {\tt paradise\_lost} Landslide will not preempt in a function named {\tt critical\_section()},
		which the test case uses to protect an internal counter used to detect the bug;
		and it will not preempt in any of the {\tt thr\_*()} thread library API functions
		for tests intended to target just the concurrency primitives.
		In future work this could be improved as annotations to be placed inside the test case code itself.
	\item Landslide finds ad-hoc synchronization patterns which students often open-code,
		rather than using the prescribed synchronization API,
		such as {\tt while (!flag) yield();} or {\tt while (xchg(...)) continue;},
		and treats them as synchronization points as described in \sect{\ref{sec:landslide-blocking}}.
	\item Landslide finds ``too suspicious'' spinwait-loops in mutex implementations
		which are neither yield- nor {\tt xchg}-loops (as described above),
		which would ordinarily be classified as infinite loop bugs,
		and reports them with a suggestive message ({\tt undesirable\_loop\_html()} in {\tt landslide.c})
		referring the student to the appropriate lecture material
		\cite{synchronization-2}.
	\item The {\tt landslide} wrapper script logs the time and command-line options of invocation
		and captures a snapshot of the student code and results of the test and saves them to AFS after each run.
		% HURDLE_VIOLATION -- no, this is for p3 (context switcher) features only
\end{itemize}

@@

\subsubsection{Test case list}

@@

@@

@@

As long as an educational user study is run on a volunteer basis,
one cannot completely avoid selection bias:
those with enough free time to participate are more likely to be the most capable students already,
who are least in need of the extra debugging help.
This was especially evident in the annotation-required P3 study from \cite{landslide},
in which only 5 groups volunteered (15\% of the 34 total who submitted P3 that semester),
among which 2 had enough after just the annotation phase and did not continue to do any in-depth testing.
In contrast, since switching to the automatically-annotatable P2,
the participation rate rose to 58\% (Table~\ref{tab:photo-of-ze-studence})
among all P2-submitting groups.
Reaching over half the class could alone be seen as a major step in mitigating said selection bias.

@@

% TODO: is this ok to report without IRB approval? ask dave
I took no special measures to compensate for bias in gender, race, or being non-natively English-speaking
during recruitment or the survey.
Surveying students to measure any differences in these demographics between study participants and the overall class population
would have required mandatory survey participation,
and in turn, a more rigorous IRB approval process.
According only to my memory of students who attended the Landslide clinics
(\sect{\ref{sec:education-pebbles-recruiting}}),
the racial diversity was roughly representative of the class at large,
and the proportion of women was in fact somewhat higher than the overall gender ratio.
Any more scientific analysis of such statistics was deemed beyond the study's scope for now.

>>>>

As described in \sect{\ref{sec:landslide-setup}},
all setup from the user's point of view is handled through the {\tt p2-setup.sh} script%
\footnote{PSU's version is called {\tt psu-setup.sh};
in this section {\tt p2-setup.sh} refers to both unless otherwise noted.
}.
It, its helper scripts (\sect{\ref{sec:landslide-glue}}),
and the {\tt landslide} script itself contain several checks to prevent
studence
from accidentally misusing Landslide in ways that could produce mysterious crashes, false bug reports, and so on
(the need for each one, as the reader might imagine, discovered through bitter experience).
These include:
\begin{itemize}
	\item {\tt p2-setup.sh} checks if the directory argument correctly points at the top-level P2 basecode directory
		rather than any subdirectories such as {\tt user/libthread/}.
	\item {\tt check-need-p2-setup-again.sh} checks if any source files in the original P2 source directory
		(the argument supplied to {\tt p2-setup.sh}),
		in case the student hoped to fix some bug and verify their fix but forgot to re-run the setup script.
	\item {\tt landslide} checks the supplied test name matches one of the endorsed Landslide-friendly tests
		(students love trying to run Landslide with {\tt racer}, {\tt largetest},
		or even the string {\tt OPTIONS}).
	\item {\tt landslide} checks if any other instance of itself is simultaneously running in the same directory,
		and if so, refuses to do so and advises the student
		to {\tt git clone} the repository afresh for simultaneous use%
		\footnote{This is ironically implemented with a non-atomic lock file
		and should really be using {\tt flock} instead.
		}.
\end{itemize}
\vspace{1em}

\noindent Landslide also includes several P2-specific instrumentations and features to cope with various student irregularities:
\begin{itemize}
	\item Quicksand emits different combinations of {\tt within\_function}/{\tt without\_function} directives
		for Landslide depending on the name of the test.
		For example, for {\tt paradise\_\allowbreak{}lost} Landslide will not preempt in a function named {\tt critical\_section()},
		which the test case uses to protect an internal counter used to detect the bug;
		and it will not preempt in any of the {\tt thr\_*()} thread library API functions
		for tests intended to target just the concurrency primitives.
		In future work this could be improved as annotations to be placed inside the test case code itself.
	\item Landslide finds ad-hoc synchronization patterns,
		such as {\tt while (!flag) yield()} or {\tt while (xchg(...)) continue},
		which students often open-code rather than using the prescribed synchronization API,
		and treats them as synchronization points as described in \sect{\ref{sec:landslide-blocking}}.
	\item Landslide finds ``too suspicious'' spinwait-loops in mutex implementations
		which are neither yield- nor {\tt xchg}-loops (as described above),
		which would ordinarily be classified as infinite loop bugs,
		and reports them with a suggestive message ({\tt undesirable\_\allowbreak{}loop\_html()} in {\tt landslide.c})
		referring the student to the appropriate lecture material
		\cite{synchronization-2}.
	\item The {\tt landslide} wrapper script logs the time and command-line options of invocation
		and captures a snapshot of the student code and results of the test and saves them to AFS after each run.
		% HURDLE_VIOLATION -- no, this is for p3 (context switcher) features only
\end{itemize}

@@

%\subsubsection{Test case list}

@@

Figure~\ref{Fig:photo-of-ze-studence} shows some resulce

\begin{figure}[t]
	\includegraphics[width=\textwidth]{photo-of-ze-studence.pdf}
	\caption{sadfasdfa}
	\label{fig:photo-of-ze-studence}
\end{figure}

@@

% probably just better to state this in one line of prose
%\begin{table}[h]
%	\begin{center}
%		\begin{tabular}{r|crcrc|r}
%			& \multicolumn{5}{c|}{\bf Did survey?} & \\
%			\bf Semester & & \bf yes & & \bf no & & \bf Total \\
%			\hline
%			CMU F'17	& &  7 & &  7 & & 14 \\
%			U. Chicago F'17	& &  1 & &  3 & &  4 \\
%			CMU S'18	& &  5 & &  5 & & 10 \\
%			PSU S'18	& & 11 & & 27 & & 38 \\
%		\end{tabular}
%	\end{center}
%	\caption{Survey response rate among students who used Landslide. (The survey was first introduced in F'17.)}
%	\label{tab:photo-of-ze-surveys}
%\end{table}

@@

As long as an educational user study is run on a volunteer basis,
one cannot completely avoid selection bias:
those with enough free time to participate are more likely to be the most capable students already,
who are least in need of the extra debugging help.
This was especially evident in the annotation-required P3 study from \cite{landslide},
in which only 5 groups volunteered (15\% of the 34 total who submitted P3 that semester),
among which 2 had enough after just the annotation phase and did not continue to do any in-depth testing.
In contrast, since switching to the automatically-annotatable P2,
the participation rate rose to 58\% (Table~\ref{tab:photo-of-ze-studence})
among all P2-submitting groups.
Reaching over half the class could already be seen as a major step in mitigating said selection bias.

@@

% TODO: is this ok to report without IRB approval? ask dave
I took no special measures to compensate for bias in gender, race, or being non-natively English-speaking
during recruitment or the survey.
Surveying students to measure any differences in these demographics between study participants and the overall class population
would have required mandatory survey participation,
and in turn, a more rigorous IRB approval process.
According only to my memory of students who attended the Landslide clinics
(\sect{\ref{sec:education-pebbles-recruiting}}),
the racial diversity was roughly representative of the class at large,
and the proportion of women I perceived was in fact somewhat higher than the overall gender ratio.
More scientific analysis of such statistics was deemed beyond the study's scope for now.

### 2018-08-19 thesis/intro.tex
education prose revise

Modern computer architectures have turned to increasing CPU core count, rather than clock speed, to improve processing power \cite{mooreslaw}.
To take advantage of multiple cores for performance, programmers must write software to execute {\em concurrently} --
using multiple {\em threads} which execute multiple parts of a program's logic simultaneously.
However, when threads access the same shared data, they may interleave in unexpected ways which change the outcome of their execution.
When an unexpected interleaving produces undesirable program behaviour,
for example, by corrupting shared data structures,
we call it a {\em concurrency bug}.
Concurrency bugs are notoriously hard for programmers to find and debug
because the specific thread interleaving required to trigger them arises at random during normal execution,
and often with very low probability.
%Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
% TODO(LAYMAN): give example of trying to open car door at same time as friend turns key to unlock it.

@@

Pronoun use will vary between more specific and more ambiguous to convey additional nuance.
The singular ``I'' is associated with my own research contributions,
while the royal ``we'' should be taken to include the reader,
such as when surveying background material or related work,
to which the author and reader share more similar relationships.
The impersonal ``the programmer'' will be referred to as she/her
to distinguish her from the thesis audience concerned with the underlying research,
and also to challenge readers' unconscious bias about gender in computer science.
Individual students who participated in the user studies will be given the more inclusive they/them.
Gender-neutral pronouns are also to be used on the author themself.

>>>>

% TODO: god.........
Modern computer architectures have turned to increasing CPU core count, rather than clock speed, to improve processing power \cite{mooreslaw}.
To take advantage of multiple cores for performance, programmers must write software to execute {\em concurrently} --
using multiple {\em threads} which execute multiple parts of a program's logic simultaneously.
However, when threads access the same shared data, they may interleave in unexpected ways which change the outcome of their execution.
When an unexpected interleaving produces undesirable program behaviour,
for example, by corrupting shared data structures,
we call it a {\em concurrency bug}.
Concurrency bugs are notoriously hard for programmers to find and debug
because the specific thread interleaving required to trigger them arises at random during normal execution,
and often with very low probability.
%Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
% TODO(LAYMAN): give example of trying to open car door at same time as friend turns key to unlock it.

@@

Pronoun use will vary between more specific and more ambiguous to convey additional nuance.
The singular ``I'' is associated with my own research contributions,
while the royal ``we'' should be taken to include the reader,
such as when surveying background material or related work,
to which the author and reader share more similar relationships.
The impersonal ``the programmer'' will be referred to as she/her
to highlight her role as the intended user, separate from the underlying research,
%to distinguish her from the thesis audience concerned with the underlying research,
and also to challenge readers' unconscious bias about gender in computer science.
Individual students who participated in the user studies will be given the more inclusive they/them.
Gender-neutral pronouns are also to be used on the author themself.
